\texttt{Taxonomic supertree construction with}\texttt{\emph{ Incertae sedis}}\texttt{ taxa}

### 2018-06-17 doc/incertae_sedis.tex

%\newcommand{\mthcomment}[1]{{\color{red} #1}\xspace}
\newcommand{\mthcomment}[1]{}

%\newcommand{\bdrcomment}[1]{{\color{cyan} #1}\xspace}
\newcommand{\bdrcomment}[1]{}

@@

In this chapter we haveve described a method for handling \emph{incertae sedis} taxa
when constructing taxonomic supertrees.
Our method incorporates the uncertain
placement information from such taxa when constructing the supertree.
It also
handles higher taxon names and allows input trees to revise the taxonomy by
determining where \emph{incertae sedis} taxa should be placed.

However, our method leaves a number of problems for future work.
One of these
problems is that we consider taxa with an \emph{incertae sedis} label to 
``invade'' a subtrees rooted by any of the taxon's siblings.  
However,
the taxonomist who made the \emph{incertae sedis} designation,
may only consider it plausible for the taxon to intrude only within specific
siblings.
Our method also assumes that an \emph{incertae sedis} label on a taxon
indicates that a taxon may attach at any depth within its siblings.  In practice, an
\emph{incertae sedis} genus with sister families (for example) is unlikely to attach
so deeply within a sister family that it becomes nested within another genus.

>>>>

\newcommand{\mthcomment}[1]{{\color{red} #1}\xspace}
\renewcommand{\mthcomment}[1]{}

\newcommand{\bdrcomment}[1]{{\color{cyan} #1}\xspace}
\renewcommand{\bdrcomment}[1]{}

@@

This paper introduces the necessary concepts and methods for handling \emph{incertae sedis}
taxa in taxonomic supertrees.
We describe the goals of preserving internal edges,
preserving higher taxon names, and preserving the opportunity to revise the taxonomy
based on placement of \emph{incertae sedis} taxa.
In addition to explaining why
the naive approach to taxonomic supertrees that ignores \emph{incertae sedis}
 taxa sacrifices these three goals, we introduce a simple and satisfying way to achieve these goals
through defining rooted splits for taxonomy edges with reduced exclude sets.
We also
provide a simple and efficient recursion for computing these splits.

Despite the simplicity of our approach, some complexities and conceptual difficulties
remain.
While the correspondence of nodes and names is a simple one-to-one and onto
relationship without \emph{incertae sedis} taxa, when \emph{incertae sedis} taxa are
introduced, the relationship might be neither one-to-one (Figures \ref{fig:Multiple-names-can},\ref{fig:interdigitation}) nor onto (Figure \ref{fig:One-name-can}).
This necessitates putting more thought into the semantics of taxon names, including
whether names might become synonyms.  This may be considered a ``complexity''.

In addition, our attempt to define reduced exclude sets for taxonomy splits yields
a semantics that is not the only concievable semantics for \emph{incertae sedis}
taxa (Figure \ref{fig:Placing-one-incertae}).  It allows \emph{incertae sedis} siblings to inter-digitate
(Figure \ref{fig:interdigitation}), which might be a surprising result to the taxonomist.  Furthermore,
in what might seem a lack of consistency, it does \emph{not} allow \emph{incertae sedis}
taxa to intrude into each other if they become siblings through the result of
placement, as is the case for taxa W and V in Figure \ref{fig:Placing-one-incertae}c.
While each of these problems can be solved separately by adjusting the definition of the
reduces exclude sets, no solution seems to be best in all respects.  This may be a result
of the fact that split-based approaches to supertrees seek a simultaneous solution to a set
of constraints.  This simultaneity can be contrasted with a solution that is phrased
as a set of sequential operations to the taxonomy, as would be the case when (for example)
(i) placing V within WX so that it is sister to W, (ii) marking V as no longer \emph{incertae sedis}, and then (iii)
placing W within V (which is no longer \emph{incertae sedis}).  This maybe considered a ``conceptual difficulty''.

Another issue that we have postponed for future work is how to more narrowly delimit
the ranges in which \emph{incertae sedis} may attach.  The current paper allows
\emph{incertae sedis} to attach anywhere within their sisters, but the taxonomist
may only consider it plausible for the taxon to intrude only within specific
siblings.  Furthermore, our method also assumes that an \emph{incertae sedis} label on a taxon
indicates that a taxon may attach at any depth within its siblings.  In practice, an
\emph{incertae sedis} genus with sister families (for example) is unlikely to attach
so deeply within a sister family that it becomes nested within another genus.
